DATA 624 Homework 9

### 2020-04-17 DATA624/HW9.Rmd
More work on HW9

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
library(gbm)
library(randomForest)
library(caret)
library(party)
library(mlbench)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

@@

How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8,24?

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
library(gbm)
library(randomForest)
library(caret)
library(party)
library(rpart)
library(mlbench)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

***I will create a simulated dataset generated by a non-linear function.  I will then train the tree based models with varying levels of pruning.  I will then look at the MSE on the training set and test set in relation to the complexity of the tree.**

```{r}
set.seed(42)

n_sample <- 350

nonlinear_function <- function(x){
  sin(1.25 * x) + 2 * cos(.25*x)
}

x <- runif(n_sample, 1, 25)
f_of_x <- nonlinear_function(x)
noise <- rnorm(n_sample, 0, 2)
y <- f_of_x + noise

df <- data.frame(y=y, x=x)
in_train <- createDataPartition(df$y, p = .8, list = FALSE, times = 1)
train_df <- df[in_train,]
test_df <- df[-in_train,]

results <- data.frame(Granularity = c(NA), MSE = c(NA), data = c(NA)) %>% na.omit()

get_mse <- function(model, data){
  y_hat <- predict(model, data)
  mse <- mean((y_hat - data$y)^2)
  return(mse)
}

for(depth in seq(1:10)){
  rtree_model <- rpart(y ~ x, data = train_df, control=rpart.control(maxdepth=depth))
  results <- rbind(results, data.frame(Granularity = depth, MSE = get_mse(rtree_model, train_df), data = "Training"))
  results <- rbind(results, data.frame(Granularity = depth, MSE = get_mse(rtree_model, test_df), data = "Test"))
}

ggplot(results, aes(Granularity, MSE, color = data, group = data)) +
  geom_line() +
  geom_point() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

**One can see that as the granularity of the tree model increases, the MSE on the training set decreases.  However the MSE on the test set initially begins to decline then increases again as the model starts to overfit the training data.**

@@

How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8.24?
